How Unprecedented Was the February 2021 Texas Cold Snap?

### 2021-12-07 doc/agu21/aggregate_rt.tex
clarify what return periods are / mean

\begin{block}{Spatially Aggregated Temperature Extremes}
    Analyzing our ``inferred heating demand per capita'' (see Methods and Data) informs the question ``what would the aggregate demand for heating have been had historic cold snaps occurred with 2020's population?''
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{ERCOT_HDD_IDF_MLE_popweighted.pdf}\\
            \caption{
                \textbf{The inferred heating demand per capita induced by the February 2021 cold snap is not unprecedented.}
                (a): time series of annual maximum inferred heating demand per capita.
                (b): the intensity-duration-frequency intervals estimated using 1950-2020 data (i.e., not using the 2021 event), overlaid by the annual maxima from the 1989, 2011, and 2021 events.
                Gray dashed lines indicate 2, 5, 10, 25, 50, and 100 year return levels.
            }\label{fig:idf_weighted}
        \end{figure}
    \end{framed}
\end{block}

>>>>

\begin{block}{Spatially Aggregated Temperature Extremes}
    We analyze \emph{inferred heating demand per capita} to understand the potential impacts of historic cold snaps if they occurred with today's population.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{ERCOT_HDD_IDF_MLE_popweighted.pdf}\\
            \caption{
                \textbf{The \emph{inferred heating demand per capita} induced by the February 2021 cold snap is not unprecedented.}
                (a): time series of annual maximum \emph{inferred heating demand per capita}.
                (b): the intensity-duration-frequency (IDF) intervals, estimated using 1950-2020 data, overlaid by annual maxima from 1989, 2011, and 2021.
                Each dashed line shows a single return level (2, 5, 10, 25, 50, 100 year) for \emph{inferred heating demand per capita} averaged over a given duration ($x$-axis).
                Gray dashed lines indicate 2, 5, 10, 25, 50, and 100 year return levels.
            }\label{fig:idf_weighted}
        \end{figure}
    \end{framed}
\end{block}

### 2021-12-07 doc/agu21/inferred-heating.tex
clarify what return periods are / mean

\begin{block}{Methods and Data}
    Engineers often use ``heating degree days'' to quantify the effect of cold temperature on people and buildings.
    We develop a spatially aggregated time series, which has the straightforward interpretation as the average heating demand experienced by a Texas resident, called ``\textbf{inferred heating demand per capita}.''
    \begin{enumerate}
        \item Take gridded (\SI{0.25}{\degree}) hourly temperature data from ERA-5 reanalysis \cite{hersbach_era5:2020} (validated using other datasets -- see online supporting information)
        \item Gridded 2020 population density from GPWv4 dataset \cite{ciesin_gpwv4:2016}
        \item For each hour $t$ and grid cell $s$: calculate
              $\text{HD}_{s,t} = \max (\SI{65}{\degree F} - T_{s,t}, \SI{0}{\degree F})$
        \item Average over the region served by the Texas Interconnection (\cref{fig:local_era5}a), weighting each grid cell by population density
    \end{enumerate}
    Finally, we compute return periods using maximum likelihood GEV models (validated using other methods -- see online supporting information).
\end{block}

>>>>

\begin{block}{Methods and Data}
    Engineers often use ``heating degree days'' to quantify the effect of cold temperature on people and buildings.
    We develop a spatially aggregated time series, which has the straightforward interpretation as the average heating demand experienced by a Texas resident, called ``\textit{\textbf{inferred heating demand per capita}}.''
    \begin{enumerate}
        \item Gridded (\SI{0.25}{\degree}) hourly temperature data from ERA-5 reanalysis \cite{hersbach_era5:2020} (validated using other datasets -- see online supporting information)
        \item Gridded 2020 population density from GPWv4 dataset \cite{ciesin_gpwv4:2016}
        \item For each hour $t$ and grid cell $s$: calculate
              $\text{HD}_{s,t} = \max (\SI{65}{\degree F} - T_{s,t}, \SI{0}{\degree F})$
        \item Average over the region served by the Texas Interconnection (\cref{fig:local_era5}a), weighting each grid cell by population density
    \end{enumerate}
    Finally, we compute return periods using maximum likelihood GEV models (validated using other methods -- see online supporting information).
\end{block}
